Numerical investigation of viscous flow instabilities through heterogeneous porous media. Advantages and limitations of unstructured and adaptive meshes in 2D and 3D domains under different mobility scenarios.

### 2017-08-11 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
Updates on Section 4.5

\medskip
In Section~\ref{section:results_homo_hete}, the impact of viscosity ratio on the growth of fingers and the triggering effect of heterogeneity on instability's nucleation were demonstrated in modified simulated Hele-Shaw cells. In order to capture the continuous development of viscous instabilities, mesh grids with sufficient resolution need to be applied over the interface between fluids. Traditional numerical methods often rely on fixed mesh with sufficient resolution throughout the domain to capture specific flow dynamics (\eg fluid instabilities, flow recirculation, heat and mass transfers etc), however computational resources may make it prohibitive for complex geometries and highly heterogeneous properties.

\medskip
Mesh adaptivity methods have been extensively used by the computational geofluid dynamic community to help capturing detailed flow dynamics, compositional non-equilibrium fluid displacement and solid-fluid interactions \citep{pluszny_2007,pietro_2014,su_2016,melnikova_2016}. In these methods, the mesh grid is continuously modified (\ie adjustments of the number and distribution of the degrees of freedom to reduce solution error) to focus resolution where is necessary as the simulation evolves in time and space. There are four main families of mesh adaptivity methods: adaptive mesh refinement (AMR), edge and face element manipulations whereas keeping the order of the element basis function fixed (h-adaptivity), mesh deformation (r-adaptivity) and changes of element basis function order (p-adaptivity). Detailed description and analysis of mesh adaptivity methods are beyond the scope of this manuscript but can be readily found in \citet{lo_book} \citep[see also][]{plewa_book,frey_book}.

@@

\medskip 
The progress of the saturation front and formation and growth of fingers in time and space can be seen in more details in the isosurface shown in Fig.~\ref{fig:3DChannel_Isosurf}. However due to the relatively low mesh resolution, in particular above and below the central channel region,

>>>>

\medskip
In Section~\ref{section:results_homo_hete}, the impact of viscosity ratio on the growth of fingers and the triggering effect of heterogeneity on instability's nucleation were demonstrated in modified simulated Hele-Shaw cells. In order to capture the continuous development of viscous instabilities, mesh grids with sufficient resolution need to be applied over the interface between fluids. Figure XXX shows a numerical simulation performed with the same geometry, boundary and initial conditions as the one shown in Fig.~\ref{fig:HeleShawHeter_VR150}, but with noticeable lower resolution (1768 instead of 26313 \PN[1]{2} element-pairs). As it can be seen \red{(COMPLETE THESE PARAGRAPH WITH BRIEF COMMENTS ON THE ONGOING SIMULATION WRT LOW MESH RESOLUTION AND ITS IMPACT ON CAPTURING FLOW DYNAMICS AND FINGER FORMATION AND GROWTH IN PARTICULAR)}.

\medskip
Traditional numerical methods often rely on fixed mesh with sufficient resolution throughout the domain to capture specific flow dynamics (\eg fluid instabilities, flow recirculation, heat and mass transfers etc), however computational resources may make it prohibitive for complex geometries and highly heterogeneous properties.  Mesh adaptivity methods have been extensively used by the computational geofluid dynamic community to help capturing detailed flow dynamics, compositional non-equilibrium fluid displacement and solid-fluid interactions \citep{pluszny_2007,pietro_2014,su_2016,melnikova_2016}. In these methods, the mesh grid is continuously modified (\ie adjustments of the number and distribution of the degrees of freedom to reduce solution error) to focus resolution where is necessary as the simulation evolves in time and space. There are four main families of mesh adaptivity methods: adaptive mesh refinement (AMR), edge and face element manipulations whereas keeping the order of the element basis function fixed (h-adaptivity), mesh deformation (r-adaptivity) and changes of element basis function order (p-adaptivity). Detailed description and analysis of mesh adaptivity methods are beyond the scope of this manuscript but can be readily found in \citet{lo_book} \citep[see also][]{plewa_book,frey_book}.

@@

\medskip 
The progress of the saturation front and formation and growth of fingers in time and space can be seen in more details in isosurfaces shown in Fig.~\ref{fig:3DChannel_Isosurf}. However due to relatively low mesh resolution, in particular above and below the central channel region, capture of fingers (formation and growth) was quite limited
